Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-01-27 Documents/M1/dgM1Curvilinear.tex
Additions to the text.  Figures for Homogeneous sphere and Riemann problem

@@

@@

To test the ability of the DG method to capture discontinuities, we solve the Riemann problem presented by \citet{OHF2012}.  
To compare with their results, we use the Eddington factor due to Levermore \citep{levermore_1984}
\begin{equation}
  \psi(h)=\f{1}{3}\big(\,5-2\,\sqrt{4-3\,h^{2}}\,\big).  
\end{equation}
The computational domain extends from $x=-0.05$ to $x=0.1$, and a discontinuity is located at $x_{\mbox{\tiny d}}=0.0$.  
The initial condition is given by
\begin{equation}
  \bcM(x,t=0)=
  \left\{
  \begin{array}{l}
  \bcM_{\mbox{\tiny L}}=\big(1,0.9999,0,0\big)^{T} \quad\text{if}\quad x\le x_{\mbox{\tiny d}} \\
  \bcM_{\mbox{\tiny R}}=\big(0.5,0,0,0\big)^{T} \quad\text{otherwise}.
  \end{array}
  \right.
\end{equation}

@@

>>>>

For time integration we employ the method lines.  
Having specified the spatial discretization for the advection terms, we consider the discretized moment equations in Eq.~\eqref{eq:semidiscreteDiscretized} as a system of ordinary differential equations (ODEs) to evolve forward in time, and write
\begin{equation}
  d_{t}\bcM_{\bk}
  =\bcF_{\bx}(\bcM)_{\bk}+\bcC(\bcM)_{\bk},
  \label{eq:momentEquationsODEs}
\end{equation}
where the operator $\bcF_{\bx}:\bbR^{d+1}\to\bbR^{d+1}$ represents the discretization of the spatial divergence operator.

@@

When solving the moment equations, we need to prevent non-physical states from developing.  
In particular, from the definition of the moments in Eq.~\eqref{eq:angularMoments}, it is easy to show that a positive distribution function implies a positive density $\cJ>0$ and a limited flux density $\cJ-|\bcH|>0$.  
Failure in maintaining these bounds can lead to an ill-posed closure problem and loss of hyperbolicity for the moment equations.

\section{Neutrino Opacities}

@@

To test the ability of the DG method to capture discontinuities, we solve the Riemann problem presented by \citet{OHF2012}.  
To compare with their results, we use the Eddington factor due to Levermore \citep{levermore_1984}
\begin{equation}
  \psi(h)=\f{1}{3}\big(\,5-2\,\sqrt{4-3\,h^{2}}\,\big).  
  \label{eq:eddingtonFactorLevermore}
\end{equation}
The computational domain extends from $x=-0.05$ to $x=0.1$, and a discontinuity is located at $x_{\mbox{\tiny d}}=0.0$.  
The initial condition is given by
\begin{equation}
  \bcM(x,t=0)=
  \left\{
  \begin{array}{l}
  \bcM_{\mbox{\tiny L}}=\big(1,0.9999,0,0\big)^{T} \quad\text{if}\quad x\le x_{\mbox{\tiny d}} \\
  \bcM_{\mbox{\tiny R}}=\big(0.5,0,0,0\big)^{T} \quad\text{otherwise}.
  \end{array}
  \right.
  \label{eq:riemannProblem}
\end{equation}
Results for $t=0.1$ are shown in Figure~\ref{fig:riemannProblem}.  
The discontinuities remain sharp, and oscillations around the discontinuities are controlled (in part due to limiting of characteristic variables; cf. Section~\ref{sec:slopeLimiters}).

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.5]{./Figures/RiemannProblem1D}
  \end{center}
  \caption{Results from the one-dimensional Riemann problem in Eq.~\eqref{eq:riemannProblem}, obtained with the DG(2)+RK3 method using $240$ elements, the HLL Riemann solver, and the Eddington factor in Eq.~\eqref{eq:eddingtonFactorLevermore}.  
  The number density $\cJ$ (black) and the number flux density $\cH_{r}$ (red) are plotted for $t=0.1$.}
  \label{fig:riemannProblem}
\end{figure}

@@

The homogeneous sphere test \citep[cf.][]{smit_etal_1997} considers of a sphere with radius $R$.  
Inside the sphere ($r<R$), the absorption opacity and equilibrium distribution function are set to constant values $\chi=\chi_{0}$ and $f_{0}=1$, respectively.  
Outside the sphere, the absorption opacity is zero.  
The steady state solution, obtained by solving the transport equation in spherical symetry, is given by
\begin{equation}
  f_{\mbox{\tiny A}}(r,\mu)=f_{0}\,\big(1-e^{-\chi_{0}\,s(r,\mu)}\big),
  \label{distributionHomogeneousSphere}
\end{equation}
where
\begin{equation}
  s(r,\mu)
  =\left\{
  \begin{array}{lll}
    r\,\mu+R\,g(r,\mu) & \mbox{if}\quad r<R, & \mu\in[-1,+1], \\
    2\,R\,g(r,\mu) & \mbox{if}\quad r \ge R, & \mu\in[(1-(R/r)^{2})^{1/2},+1], \\
    0 & \mbox{otherwise},
  \end{array}
  \right.
\end{equation}
and $g(r,\mu)=[1-(r/R)^{2}(1-\mu^{2})]^{1/2}$.

Similar to \citet{oConnor_2015}, we solve two versions of this test.  
One where the absorption opacity is set to $\chi_{0}=10^{-4}$~cm$^{-1}$ (thick; similar to the electron capture opacity encountered in the center of a collapsed stellar core), and one where the absorption opacity is set to $\chi_{0}=10^{-6}$~cm$^{-1}$ (thin).  
Both simulations are run with a computational domain with $r\in[0,500]$~km using 100 uniform elements $\Delta r=5$~km.  
Thus, in the thick case, the Knudsen number is small, $\mbox{Kn}=(\chi_{0}\,\Delta r)^{-1}=0.02$, while it exceeds unity in the thin case, $\mbox{Kn}=2$.  
The radius of the sphere is set to $R=100$~km.

Results are plotted in Figure~\ref{fig:homogeneousSphere1D}.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.32]{./Figures/HomogeneousSphere1D_Chi_1e-4_Density} &
      \includegraphics[scale=0.32]{./Figures/HomogeneousSphere1D_Chi_1e-6_Density} \\
      \includegraphics[scale=0.32]{./Figures/HomogeneousSphere1D_Chi_1e-4_FluxFactor} &
      \includegraphics[scale=0.32]{./Figures/HomogeneousSphere1D_Chi_1e-6_FluxFactor} \\
      \includegraphics[scale=0.32]{./Figures/HomogeneousSphere1D_Chi_1e-4_EddingtonFactor} &
      \includegraphics[scale=0.32]{./Figures/HomogeneousSphere1D_Chi_1e-6_EddingtonFactor}
    \end{tabular}
  \end{center}
  \caption{Results from solving the Homogeneous sphere problem with the DG(2)+SIRK2 method using $100$ elements.  
  In the left panels, the absorption opacity is set to $\chi_{0}=10^{-4}$~cm$^{-1}$, while in the right panels it is set to $\chi_{0}=10^{-6}$~cm$^{-1}$.  
  The numerical results (dashed lines) are compared to the analytical solution (solid lines).}
  \label{fig:homogeneousSphere1D}
\end{figure}

\subsection{Milne's Problem}

\section{Mini-app: Transport Problems with Tabulated Neutrino Opacities}
